---
title: "Data Science for Beginners, University of Essex"
subtitle: "Day 4: Working with Data Frame"
author: "Dr. Howard Liu"  #If multiple, 'c("A", "B")'
date: "13-01-2022"  #r Sys.Date()
output:
  html_document: default
  pdf_document: default
  nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Learning Objectives Today
1. Import and export data files
2. Explore the data 
3. Pipe operator
4. Working with data frames: dplyr functions for data wrangling


# 1. Import data
To read in (=import) a data set from a file, we use read.XXX functions, such as

* read.csv        Comma-separated text file
* read.table      Tab-separated text file
* read.dta        Stata file
* read.spss       SPSS file
* read.xlsx       Excel file

To load data, your computer needs to know where the data file is saved. The key is to set the path for R to locate

```{r, include=T, message=F, warning=FALSE}
myPath <- "/Users/howardliu/Dropbox/Essex/data-programming-beginners/Lecture/" # THIS IS IMPORTANT!
setwd(myPath)

library(tidyverse) # I need this package for the as_tibble function. I will introduce tidyverse this inclusive package more later on
world.data = read.csv("world.csv") %>% as_tibble() # here I make it a tibble data frame just for illustration purpose
```

## 1-2 Export data
To export data, use the write.csv() function
```{r, include=T, message=F, warning=FALSE}
myPath <- "/Users/howardliu/Dropbox/Essex/data-programming-beginners/Lecture/"
setwd(myPath) # set working directory will let R to use this as your default path

# write.csv(world.data, file = "world_tmp.csv") # You can export your data as a csv file. 

# save(world.data, world.data2, file ="world_tmp.rda") # This allows you to save multiple data frames together into one file
```



# 2. Explore your data
Whenever you read in a data set from a file, it's always a good idea to take a look at it first to have an idea about what it looks like. There are a few functions that will let us take a look at the data set.

The dim function tells us the dimension (the number of rows and the number of columns).
```{r, include=T, message=F, warning=FALSE}
dim(world.data)
```
So this contains 191 rows (observations) and 62 columns (variables)


The `head()` and `tail()` functions let us see the first or last 5 rows of the data set.
```{r, include=T, message=F, warning=FALSE}
head(world.data)

tail(world.data)
```

Or we can do this using the square brackets:

```{r, include=T, message=F, warning=FALSE}
world.data[1:10, 1:5]
```

This tells R to show the first 10 rows and first 5 columns. We can easily see that country is the unit of observation. We can also see that there are variables such as colony, confidence, decentralization, and dem_other. 

When working with a large data set like this one, you may want to  use the spreadsheet view (just like Excel). To do so, we use the `View()` function, which opens a spreadsheet tab in the script pane. 

```{r, include=T, message=F, warning=FALSE}
# View(world.data)
```

If you want to take a further look at a specific variable, you can use the `$` to call it and the `summary()` function. 
```{r, include=T, message=F, warning=FALSE}
world.data$colony
summary(world.data$colony)
```

Or you can simply summarize the entire data frame, which in some cases might be useful.
```{r, include=T, message=F, warning=FALSE}
summary(world.data)
```


# 3. Understanding pipe operator %>% in R
Before introducing the dplyr package (a fast, powerful tool for working with data frame), we need to know the pipe operator first. 
```{r, message=F, warning=FALSE}
# The pipe operator makes it possible to easily chain a sequence of calculations.

# First, install it from this package (You don't have to if you have installed the tidyverse package)
library(magrittr)

# Suppose we create an object and assign values. If we want to do multiple calculations, we need to write a lot of parentheses, which can be very annoying. 

# For example:
x <- c(0.109, 0.359, 0.63, 0.996, 0.515, 0.142, 0.017, 0.829, 0.907)

# Like compute the logarithm of `x`, return suitably lagged and iterated differences, 
# compute the exponential function and round the result
round(exp(diff(log(x))), 1)

# Same task can be completed by using pipes. Everything looks much clearer and easy. 
x %>% log %>% diff %>% exp %>% round(., 1)
```
The hot key for %>% is `Command` + `shift` + `m` (on MAC)

Now, let's proceed on to discuss dplyr functionality then.

# 4. dplyr in R
It is a fast, powerful tool for working with data frame. You don't need to worry about parentheses when writing commands anymore. It is magical! 

dplyr is based on the concepts of functions as "verbs" that manipulate data frames. It's also powerful when you manage large datasets,

Single data frame functions / verbs:

* `filter()`: pick rows matching criteria
* `slice()`: pick rows using index(es)
* `select()`: pick columns by name
* `pull()`: grab a column as a vector
* `rename()`: rename specific columns
* `arrange()`: reorder rows
* `mutate()`: add new variables
* `transmute()`: create new data frame with variables
* `distinct()`: filter for unique rows
* `sample_n()` / `sample_frac()`: randomly sample rows
* `summarise()`: reduce variables to values
* ... (many more)


## dplyr rules for functions

1. First argument is *always* a data frame

2. Subsequent arguments say what to do with that data frame

3. *Always* return a data frame

4. Don't modify in place

5. Lazy evaluation magic

## Let's load the dplyr package first.

You can either use the dplyr package directly. Or you can use the tidyverse package that includes other important packages that you’re likely to use in everyday data analyses (i.e., ggplot2, dplyr, readr, tibble, stringr etc.)

```{r, include=T, message=F, warning=FALSE}
library(dplyr) 
library(tidyverse) # or this
```

### Example Data

We will demonstrate dplyr's functionality using the `world.data` again.

### `filter()` - Show only countries in the EU
```{r message=FALSE}
world.data %>% filter(eu =="EU Member state") 
```

### `filter()` - Show countries in the EU and GDP per capita is high. Plus I only want to show the country name and not anything else.
```{r message=FALSE}
world.data %>% filter(eu =="EU Member state" & gdp_cap2 == "High") %>% select(country) 
```
### `slice()` - Pick the first 10 countries
```{r message=FALSE}
world.data %>% slice(1:10) 

```
### `slice()` - Last 5 countries and show country names only
```{r message=FALSE}
world.data %>% slice( (n()-4) : n() ) %>% select(country)
```

### `select()` - Pick Individual Columns
```{r message=FALSE}
world.data %>% select(country, eu, gdp_cap2) 
```


### `select()` - Exclude Columns
```{r message=FALSE}
world.data %>% select(-country, -eu, -gdp_cap2) 
```

### `select()` - Matching: Pick variables that contain these characters
```{r message=FALSE}
world.data %>% select(contains("gdp"), 
                   contains("dem"))
```
### `mutate()` - Modify columns and create a new variable
```{r message=FALSE}
world.data %>% select(1:2) %>% mutate(newVar = paste(country, colony, sep="/") )
```
### `rename()` - Change column names
```{r message=FALSE}
world.data %>% rename(gdp_08 = gdp08) %>% select(gdp_08)
```
### `summarise()` with `group_by()`
It will do calculation for the group that you want to create.

```{r message=FALSE}
world.data %>% group_by(eu) %>% 
  summarise(meanGDP = mean(gdp08, na.rm = TRUE)) # Here I calculate the mean. You could replace by using `sum()` for example.
```

### `left_join()` - Merge two datasets
```{r message=FALSE}
# first I create two datasets
dat1 = world.data %>% select(1:3)
dat2 = world.data %>% select(1:2, 4)

dat_merged = left_join(dat1, dat2, by = c("country" = "country", "colony" = "colony")) # merge them by picking two variables that are used as mergers
dat_merged 
```

### Great! We've finished the lecture and you can go to day4 exercise to do some additional practices for today's content.
