---
title: "R code for Data Science for Beginners"
subtitle: " Day 4: Individual Exercise"
author: "[[Your name please]]"  #If multiple, 'c("A", "B")'
date: "2024-9-12"  #r Sys.Date()
output:
  pdf_document: default
  html_document: default
---

# Clean up your workspace 
```{r}
rm(list=ls(all=TRUE)) # remove all the named objects visible in the environment 
cat("\014") # clean your console
```

## 1. Let's do more exercises with dplyr (with a different dataset) 

Please download the nycflights13 data by installing this package called `nycflights13`

```{r}
# install.packages("nycflights13")
library("nycflights13")
```

### 1-1: Please find all March flights in the data (the dataset is named "flights") flights

WRITE YOUR ANSWER (code) HERE

### 1-2 :Create a new variable as date with a format like this 1/1/2013, using the `mutate()` function

WRITE YOUR ANSWER (code) HERE

### 1-3: Change column name tailnum to tail_number

WRITE YOUR ANSWER (code) HERE

### 1-4: Group flights by their origins

WRITE YOUR ANSWER (code) HERE

### 1-5: Count how many flights departing from JFK on 2013-12-31?

WRITE YOUR ANSWER (code) HERE

### 1-6: Calculate the average hours of delay in departure for all flights from JFK

WRITE YOUR ANSWER (code) HERE

Finally, execute the entire contents of this file. Make sure that you don't get any error message. If you get an error message, it's probably because you forgot to comment out something. 


