---
title: "R code for Data Science for Beginners"
subtitle: " Day 5: Individual Exercise"
author: "[[Your name please]]"  #If multiple, 'c("A", "B")'
date: "2024-9-13"  #r Sys.Date()
output:
  pdf_document: default
  html_document: default
---

# clean up your workspace 
```{r}
rm(list=ls(all=TRUE)) # remove all the named objects visible in the environment 
cat("\014") # clean your console
```

# Load the data 
```{r}
world.data = read.csv("Data/world.csv")
```

Remember that Quarto uses a relative path so always save your data in the same folder (or under the same folder) with your Quarto code. 

## Democracy and female representation 

Do democratic countries (`democ_regime == "Yes"`) have better female representation than non-democratic countries? Please answer this question by showing some graphs to assess the relationship between these two variables. 


### Initial data clean up

Analyze y
```{r}
summary(world.data $ women09)
```

Analyze x
```{r}
summary(world.data $ democ_regime)
```

Note: There are some missing values. Labels are not intuitive. So, we will deal with these two first.


### Create a smaller data set that omits NA observations.
```{r}
women.dem <- world.data[ is.na(world.data $ women09) == FALSE & is.na(world.data $ democ_regime) == FALSE, ]
```

Let's make sure that we did this correctly. 

```{r}
summary(women.dem $ women09)
summary(women.dem $ democ_regime)
```

We can see that NA cases have been correctly removed. 


### Re-labeling the values
```{r}
women.dem $ dem.dum <- factor(women.dem $ democ_regime, 
                              levels = c("Yes", "No"),
                              labels = c("Democracy", "Autocracy"))

rm(world.data) # we don't need this anymore
```

## Q1: Describe Y (numerically and graphically) 


### Numerical summary

WRITE YOUR ANSWER (code) HERE

### Graphical summary

WRITE YOUR ANSWER (code) HERE

## Q2: Describe X (numerically and graphically) 


### Numerical summary (frequency table)

WRITE YOUR ANSWER (code) HERE

### Graphical summary (bar chart)

WRITE YOUR ANSWER (code) HERE

## Q3: Describe X-Y graphically

### Histograms

WRITE YOUR ANSWER (code) HERE

### Box-plots

WRITE YOUR ANSWER (code) HERE